---
title: "Assignment 2: Coding Basics"
author: "Samantha White-Murillo"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<Samantha White-Murillo>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 30, increasing by threes. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. 
sequence <- seq(1,30,3) #creating a sequence and labeling it

#2. 
mean(sequence) # operation with the sequence
median(sequence)

#3. 
mean(sequence) > median(sequence) #comparison of results

```

## Basics, Part 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r basics 2}
#Student
names_of_students <- c("Samantha","Ivy","Juan")

#Student_scores
score_of_students <- seq(30,100,length.out = 3)

#Passed_the_test
passed <- score_of_students > 50

#Making_data_frame
is.data.frame(names_of_students)
df_name_of_students <- is.data.frame(names_of_students)
df_name_of_students
class(df_name_of_students)

#adding_frame
df <- cbind(df_name_of_students, score_of_students, passed)
class(df)
df

#Other option of data frame creation
create_df <- data.frame(names_of_students,score_of_students,passed)
create_df

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: matrices can only contain a single class of data, while data frames can consist of many different classes of data.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r basics 3}



```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:
